---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
---

```{r include=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggalt)   

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r include = FALSE}
TB_usage = read.csv("./data/TB_expenditure_utilisation.csv")
```

```{r tidy data, include = FALSE}
table(TB_usage$g_whoregion) #We will use WPR(Western Pacific) , SEA(South East Asia)
TB_usage = TB_usage %>%
  select(
    country, year, g_whoregion,hospd_mdr_dur,
    rcvd_fld, rcvd_lab, rcvd_mdrmgt, rcvd_orsrvy, rcvd_oth, rcvd_patsup, 
    rcvd_prog, rcvd_sld, rcvd_staff, rcvd_tbhiv, rcvd_tpt
  ) %>%
  filter(
    g_whoregion == "WPR" | g_whoregion == "SEA"
  ) %>%
  select(-g_whoregion)  %>%
  rename(
    patient = rcvd_patsup,
    staff = rcvd_staff,
    programme = rcvd_mdrmgt,
    drugs = rcvd_fld,
    other = rcvd_oth
  )

colSums(is.na(TB_usage)) #remove rcvd_tpt because there are many NAs.

TB_usage = TB_usage %>% 
  select(-rcvd_tpt) 

colSums(is.na(TB_usage))

TB_usage = TB_usage %>%
  na.omit(rcvd_fld)

TB_usage = TB_usage[TB_usage$country %in% names(which(table(TB_usage$country) > 2)), ] 
#remove the data that doesn't have 2017,2018,2019
```

```{r Variable:Types of Usage funding in each country in 2017-2019, include=FALSE}
unique(TB_usage$country)
colSums(TB_usage == 0)


TB_usage %>%
  select(-country, -year) %>%
  colMeans() %>%
  order() 

#top 5 usage type of fundings are: patient, staff, programme, drugs, other

usage_2017_top5 = TB_usage %>% 
  filter(year == "2017") %>%
  select(country, patient, staff, programme, drugs, other) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')

usage_2018_top5 = TB_usage %>% 
  filter(year == "2018") %>%
  select(country, patient, staff, programme, drugs, other) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')

usage_2019_top5 = TB_usage %>% 
  filter(year == "2019") %>%
  select(country, patient, staff, programme, drugs, other) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')
```


The main dataset for our project is made by  World Health Organization. It can be access at [here](https://www.who.int/teams/global-tuberculosis-programme/data). The dataset which consists types of usage funding has a range from 2017-2019, and other datset that consists estimated mortality of TB cases has a range of 2000 to 2019. To make the data consistent, we will use the data from 2017 to 2019.

Variables that we are going to use are:

* **patient**: Funding received for patient support (US Dollars)
* **staff**: Funding received for National TB Programme staff (central unit staff and subnational TB staff)  (US Dollars)
* **programme**:Funding received for programme costs to treat drug-resistant TB (US Dollars)
* **drugs**:Funding received for drugs to treat drug-susceptible TB (US Dollars)
* **other**:Funding received for all other budget line items (US Dollars)
* **hospd_mdr_dur**: Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized (days)


There are other types of usage funding, but we selected top 5 by taking the mean of each types and ordered them.
# Regression Analysis
## Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized (days)
```{r include=FALSE, fig.width = 8, fig.height = 10}
TB_duration_2017 = TB_usage %>%
  filter(year == "2017") %>%
  mutate(hospd_2017 = hospd_mdr_dur) %>%
  select(country, hospd_2017)

TB_duration_2019 = TB_usage %>%
  filter(year == "2019") %>%
  mutate(hospd_2019 = hospd_mdr_dur) %>%
  select(country, hospd_2019)

TB_duration = inner_join(TB_duration_2017, TB_duration_2019)
```
We can see that most of the country had the same or decreased estimated average duration of stay.
```{r echo=FALSE, message=FALSE}
TB_duration = inner_join(TB_duration_2017, TB_duration_2019)
TB_duration_2017_2019 =  tidyr::gather(TB_duration, year, value, -country)
TB_duration$country <- reorder(TB_duration$country, TB_duration$hospd_2019)
TB_duration_2017_2019$country <- factor(TB_duration_2017_2019$country, levels = levels(TB_duration$country))

ggplot(TB_duration, aes(y = country)) + 
  geom_point(data = TB_duration_2017_2019, aes(x = value, color = year), size = 2) +
  geom_dumbbell(aes(x = hospd_2017, xend = hospd_2019), size=3, color="#e3e2e1", 
                colour_x = "red", colour_xend = "blue",
                dot_guide=TRUE, dot_guide_size=0.25) +
  theme_bw() +
  scale_color_manual(name = "", values = c("red", "blue") ) +
  labs(x = "Estimated average duration of stay in hospitals(days)", y = "Country", 
       title = "Estimated average duration of stay if patients \n starting MDR-TB/XDR-TB
       treatment are hospitalized (days)") +
    theme(axis.text.y = element_text(angle = 45, vjust = 0.5, hjust = 1))

```

# Transformation and Assumption of the Normality 

Below is the density graph of estimated average duration of stay.
Since the graph is right-skewed, we need to make some changes to make this to satisfy the assumptions of the linear model.

```{r GGPLOT, message=FALSE, warning=FALSE, echo=FALSE}
TB_usage %>%
  ggplot(aes(x=hospd_mdr_dur)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  geom_density(alpha = 0.3, fill = "red") +
  labs(x = "Estimated average duration of stay(days)", title = "Histogram of Estimated Average Duration of Stay") 
```

We used log transformation to make the graph normal. Below is the graph after the transformation, and we can say that it is more bell-shaped than the previous graph.
```{r message=FALSE, warning=FALSE, echo=FALSE}
TB_usage %>%
  ggplot(aes(x=hospd_mdr_dur)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  geom_density(alpha = 0.3, fill = "red") +
  labs(x = "Estimated average duration of stay(days)", title = "Histogram of Estimated Average Duration of Stay") +
    scale_x_log10()

```

# Results
With the regression analysis, we can predict the Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized. We will use the top 5 usage of funding and year. 
We have included the interaction between **patient** and **staff** because there could be a possible relationship that the funding for patient support can have different duration of stay to different national TB Programme staff who received funding.

```{r include=FALSE }
TB_burden = read.csv("./data/TB_burden_countries.csv")

TB_burden = TB_burden %>%
  mutate(
    e_mort_pro = as.numeric(e_mort_num)/as.numeric(e_pop_num),
    e_mort_pro_100k = as.numeric(e_mort_num)/as.numeric(e_pop_num) * 100000
  )
df1 = TB_burden %>% 
  select(country, year, e_inc_num, e_pop_num, e_mort_pro_100k, e_mort_pro )

df2 = TB_usage %>%
  select(country, year, hospd_mdr_dur, patient,  staff,  programme, drugs, other, rcvd_prog, rcvd_sld, staff, rcvd_tbhiv )
  
TB_regression = merge(df1, df2, by=c("year")) %>%
  na.omit() %>%
  mutate(
   log_hosp = log(hospd_mdr_dur) 
  )

TB_regression = TB_regression %>% 
  filter_all(all_vars(!is.infinite(.)))

TB_usage = TB_usage %>%
  mutate(
   log_hosp = log(hospd_mdr_dur) 
  ) %>%
  filter_all(all_vars(!is.infinite(.)))

lm_model1 = lm(log_hosp ~ year + patient + staff + programme + drugs + other + patient*staff, data = TB_usage)

lm_model2 = lm(log_hosp ~ year + patient + staff + programme + drugs + other, data = TB_usage)
```

Based on transformation above, the model we are going to fit is:
$$log(average\ duration\ of\ stay\ for\ treatment) = \beta_0 + \beta_1 year + \beta_2 patient  + \beta_3 staff + \beta_4 programme + \beta_5 drugs + \beta_6 other$$

```{r echo=FALSE}
summary(lm_model1)
```

Based on the summary, we can see that some of our main effects are not significant due to small data.
## All Countries
So, we are going to look at all the countries' data and find the multiple linear regression.

```{r include=FALSE}
all_lm_model1 = lm(log_hosp ~ year + patient + staff + programme + drugs + other + patient*staff, data = TB_regression)

all_lm_model2 = lm(log_hosp ~ year + patient + staff + programme + drugs + other, data = TB_regression)
```
Looking at the regression model for all countries,  all variables are significant but the interaction between **patient** and **staff** is not significant.
Below is the ANOVA Test and we can conclude that since the pvalue is bigger than 0.05, the bigger model is not significant.
```{r}
options("scipen"=100, "digits"=4)
summary(all_lm_model2)
anova(all_lm_model1,all_lm_model2)
```

# Prediction
Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized (days) in 2020:
```{r include=FALSE}
TB_regression %>%
  na.omit()
mean(TB_regression$patient)
mean(TB_regression$staff)
mean(TB_regression$programme)
mean(TB_regression$drugs)
mean(TB_regression$other)

newdata = data.frame(year = 2020, patient = 5601390, staff = 13714771, programme = 3517401, drugs = 8813787, other = 7378395) 

```

```{r include=FALSE}
predict(lm_model1, newdata, interval="predict")
```
After putting the average funding for each variables, we are  95% confident that the in 2020,  the average stay at the hospitals will fall between **-0.3268** and **5.75** days