---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
---

```{r include=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggalt)   
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r include = FALSE}
TB_usage = read.csv("./data/TB_expenditure_utilisation.csv")
```

```{r tidy data, include = FALSE}
table(TB_usage$g_whoregion) #We will use WPR(Western Pacific) , SEA(South East Asia)
TB_usage = TB_usage %>%
  select(
    country, year, g_whoregion,hospd_mdr_dur,
    rcvd_fld, rcvd_lab, rcvd_mdrmgt, rcvd_orsrvy, rcvd_oth, rcvd_patsup, 
    rcvd_prog, rcvd_sld, rcvd_staff, rcvd_tbhiv, rcvd_tpt
  ) %>%
  filter(
    g_whoregion == "WPR" | g_whoregion == "SEA"
  ) %>%
  select(-g_whoregion)

colSums(is.na(TB_usage)) #remove rcvd_tpt because there are many NAs.

TB_usage = TB_usage %>% 
  select(-rcvd_tpt) 

colSums(is.na(TB_usage))

TB_usage = TB_usage %>%
  na.omit(rcvd_fld)

TB_usage = TB_usage[TB_usage$country %in% names(which(table(TB_usage$country) > 2)), ] 
#remove the data that doesn't have 2017,2018,2019
```

```{r Variable:Types of Usage funding in each country in 2017-2019, include=FALSE}
unique(TB_usage$country)
colSums(TB_usage == 0)

TB_usage %>%
  select(-country, -year) %>%
  colMeans() %>%
  order() 
#top 5 usage type of fundings are: rcvd_patsup, rcvd_staff, rcvd_mdrmgt, rcvd_fld, rcvd_oth

usage_2017_top5 = TB_usage %>% 
  filter(year == "2017") %>%
  select(country, rcvd_patsup, rcvd_staff, rcvd_mdrmgt, rcvd_fld, rcvd_oth) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')

usage_2018_top5 = TB_usage %>% 
  filter(year == "2018") %>%
  select(country, rcvd_patsup, rcvd_staff, rcvd_mdrmgt, rcvd_fld, rcvd_oth) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')

usage_2019_top5 = TB_usage %>% 
  filter(year == "2019") %>%
  select(country, rcvd_patsup, rcvd_staff, rcvd_mdrmgt, rcvd_fld, rcvd_oth) %>%
  melt(id.vars = 'country', variable.name = 'usage_type')
```


The main dataset for our project is made by  World Health Organization. It can be access at [here](https://www.who.int/teams/global-tuberculosis-programme/data). The dataset which consists types of usage funding has a range from 2017-2019, and other datset that consists estimated mortality of TB cases has a range of 2000 to 2019. To make the data consistent, we will use the data from 2017 to 2019.

Variables that we are going to use are:

* **rcvd_patsup**: Funding received for patient support (US Dollars)
* **rcvd_staff**: Funding received for National TB Programme staff (central unit staff and subnational TB staff)  (US Dollars)
* **rcvd_mdrmgt**:Funding received for programme costs to treat drug-resistant TB (US Dollars)
* **rcvd_fld**:Funding received for drugs to treat drug-susceptible TB (US Dollars)
* **rcvd_oth**:Funding received for all other budget line items (US Dollars)
* **hospd_mdr_dur**: Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized (days)


There are other types of usage funding, but we selected top 5 by taking the mean of each types and ordered them.

# Type of Usage Funding in Each Country
## 2017
Most countries got funded for drugs to treat drug-susceptible TB (US Dollars)
```{r echo=FALSE, fig.width = 12, fig.height = 12}
ggplot(data = usage_2017_top5, aes(x = country, y = value, fill = usage_type)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
 labs(x = "Countries", title = "Types of Usage funding in each country in 2017") +
  scale_y_continuous(name = "US Dollars($)", labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```

## 2018
The funding received for National TB Programme staff has increased than last year.
```{r echo=FALSE, fig.width = 8, fig.height = 10}
ggplot(data = usage_2018_top5, aes(x = country, y = value, fill = usage_type)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
 labs(x = "Countries", title = "Types of Usage funding in each country in 2018") +
  scale_y_continuous(name = "US Dollars($)", labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```

## 2019
Funding received for all other budget line items was decreased.
```{r echo=FALSE, fig.width = 8, fig.height = 10}
ggplot(data = usage_2019_top5, aes(x = country, y = value, fill = usage_type)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
 labs(x = "Countries", title = "Types of Usage funding in each country in 2019") +
  scale_y_continuous(name = "US Dollars($)", labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1))
```

# Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized (days)
```{r include=FALSE, fig.width = 8, fig.height = 10}
TB_duration_2017 = TB_usage %>%
  filter(year == "2017") %>%
  mutate(hospd_2017 = hospd_mdr_dur) %>%
  select(country, hospd_2017)

TB_duration_2019 = TB_usage %>%
  filter(year == "2019") %>%
  mutate(hospd_2019 = hospd_mdr_dur) %>%
  select(country, hospd_2019)

TB_duration = inner_join(TB_duration_2017, TB_duration_2019)
```
We can see that most of the country had the same or decreased estimated average duration of stay.
```{r echo=FALSE, message=FALSE}
TB_duration = inner_join(TB_duration_2017, TB_duration_2019)
TB_duration_2017_2019 =  tidyr::gather(TB_duration, year, value, -country)
TB_duration$country <- reorder(TB_duration$country, TB_duration$hospd_2019)
TB_duration_2017_2019$country <- factor(TB_duration_2017_2019$country, levels = levels(TB_duration$country))

ggplot(TB_duration, aes(y = country)) + 
  geom_point(data = TB_duration_2017_2019, aes(x = value, color = year), size = 2) +
  geom_dumbbell(aes(x = hospd_2017, xend = hospd_2019), size=3, color="#e3e2e1", 
                colour_x = "red", colour_xend = "blue",
                dot_guide=TRUE, dot_guide_size=0.25) +
  theme_bw() +
  scale_color_manual(name = "", values = c("red", "blue") ) +
  labs(x = "Estimated average duration of stay in hospitals(days)", y = "Country", 
       title = "Estimated average duration of stay if patients \n starting MDR-TB/XDR-TB
       treatment are hospitalized (days)") +
    theme(axis.text.y = element_text(angle = 45, vjust = 0.5, hjust = 1))

```

# Transformation and Assumption of the Normality 

Below is the density graph of estimated average duration of stay.
Since the graph is right-skewed, we need to make some changes to make this to satisfy the assumptions of the linear model.

```{r GGPLOT, message=FALSE, warning=FALSE, echo=FALSE}
TB_usage %>%
  ggplot(aes(x=hospd_mdr_dur)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  geom_density(alpha = 0.3, fill = "red") +
  labs(x = "Estimated average duration of stay(days)", title = "Histogram of Estimated Average Duration of Stay") 
```

We used log transformation to make the graph normal. Below is the graph after the transformation, and we can say that it is more bell-shaped than the previous graph.
```{r message=FALSE, warning=FALSE, echo=FALSE}
TB_usage %>%
  ggplot(aes(x=hospd_mdr_dur)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  geom_density(alpha = 0.3, fill = "red") +
  labs(x = "Estimated average duration of stay(days)", title = "Histogram of Estimated Average Duration of Stay") +
    scale_x_log10()

```

# Results
With the regression analysis, we can predict the Estimated average duration of stay if patients starting MDR-TB/XDR-TB treatment are hospitalized. We will use the top 5 usage of funding and year. 
We have included the interaction between **rcvd_patsup** and **rcvd_staff** because there could be a possible relationship that the funding for patient support can have different duration of stay to different national TB Programme staff who received funding.

```{r include=FALSE }
lm_model1 = lm(hospd_mdr_dur ~ year + rcvd_patsup + rcvd_staff + rcvd_mdrmgt + rcvd_fld + rcvd_oth + rcvd_patsup*rcvd_staff, data = TB_usage)

lm_model2 = lm(hospd_mdr_dur ~ year + rcvd_patsup + rcvd_staff + rcvd_mdrmgt + rcvd_fld + rcvd_oth, data = TB_usage)
```

Based on transformation above, the model we are going to fit is:
$$log(average\ duration\ of\ stay\ for\ treatment) = \beta_0 + \beta_1 year + \beta_2 patient  + \beta_3 staff + \beta_4 programme + \beta_5 drugs + \beta_6 other + \beta_7 patient*staff$$

```{r echo=FALSE}
summary(lm_model1)
```

Based on the summary, we can see that some of our main effects and the interaction term are statistically significant. 

The F-statistics is 214 with 7 and 7054df. It’s obvious that the model is statistically significant under any reasonable critical value. It also has a R^2 of 0.1751, saying that 17.51% can be explained the indicated variables. 

We are also going to test the ANOVA test for accuracy:
All the variables are less than the significant value(0.05) and we can conclude that at larger model is superior.
```{r echo=FALSE}
anova(lm_model1, lm_model2)
```